---
title: "Likelihood Totorial"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Agenda
- ~10 minutes slides
- ~20 min code
- ~10 exercise 
- ~5 minutes dicussion


```{r}
library(tidyverse)
```


# Estimating Knowledge of teachers
Suppose we made each teacher answer $n$ question and we got the following results (1 is correct, 0 is wrong).

```{r}
thoma = c(rep(1, 9), rep(0, 1))   # 9 correct 1 wrong
palle = c(rep(1, 2))              # palle has other things to do and only answered 2 question correct
iker = c(rep(1, 82), rep(0, 18))  # iker is very dilligent and answers lots of questions (82 correct 18 wrong)
```

Define a function to calculate the binomial distribution:
```{r}
# you can simply use dbinom
binomial <- function(n_correct, n_trials, p){
  choose(n_trials, n_correct) * p^n_correct * (1-p)^(n_trials-n_correct)  # what does the choose function do?
}
```

## Calculate the maximum likelihood 
```{r}
grid = seq(from=0, to=1, by=0.001)

n_correct = sum(thoma == 1)
n_trials = length(thoma)

likelihood = binomial(n_correct, n_trials, grid)

maximum_likelihood = n_correct/n_trials  # derived analytically
# or
maximum_likelihood = grid[likelihood == max(likelihood)] # argmax function
# what is meant by the argmax function?
```

## Plot the maximum likelihood
```{r}
tibble(grid, likelihood) %>% 
  ggplot(aes(grid,likelihood)) +
  geom_line(color="steelblue") + 
    # aesthetic stuff:
  geom_vline(xintercept=maximum_likelihood, linetype="dashed", alpha=0.5) + 
  annotate("point", x = maximum_likelihood, y=max(likelihood)) + 
  theme_minimal() +
  labs(x="Probability", y="Likelihood")
```


# Likelihood Ratio Test
```{r}
h0 <- log(likelihood[grid==0.5])
ha <- log(likelihood[grid==maximum_likelihood])

lik_ratio <- 2*(ha - h0)  # what does this equation correspond to in the paper?

p_value = pchisq(q = lik_ratio, df = 1, lower.tail = F)
# interpret the p-value, what does it tell you?
```

# Exercise 
- [x] go through the code and make sure you understand it
- [ ] do if for the two other (you could make it into a function)
- [ ] make a plot with all three likelihoods

---





